Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-03 paper/tv.tex
m

Here, we regard \code{k} and \code{x} are meta-variables.  \code{k^\tau} represents a constant of basic type \code{\tau}. \textbf{var} represents the countably infinite set of arbitrary distinct identifiers. 
\anyevent{a} refers to a singleton event coming from the finite set of events \code{\Sigma}. 
%Without loss of generality, any event can be further parametrized with constant values, e.g. \code{\anyevent{a}.0} and  \code{\anyevent{a}}.1 are distinct two events, leading to infinite alphabet.
We assume that programs we use are well-typed conforming to basic types \code{\tau} (we take () as the \code{void} type).
%Each events raising expression \code{\textbf{event}} is parametrized with an event, to trigger such a single event. 
A program \code{\mathcal{P}} comprises a list of 
%data type declarations \code{{datat^*}} and a list of 
method declarations \code{{meth^*}}. 
Here, we use the \code{*} superscript to denote a finite list (possibly empty) of items, for example, \code{{x^*}} refers to a list of variables, \code{x_1, ...,\ x_n}. 
%Besides, \code{d} denotes the name of a user-defined data type, \code{f} denotes a field name.

Each method \code{meth} has a name \code{mn}, an expression-oriented body \code{e}, also is associated with a precondition  \code{\effect_{pre}} and a postcondition \code{\effect_{post}} (the syntax of effects specification \code{\effect} is given in \figref{fig:Syntax_of_Types_and_Effects}).
The language allows each iterative loop to be optimized to an equivalent tail-recursive method, where mutation on parameters is made visible to the caller. The technique of translating away iterative loops is standard
%\citeref{insa2015automatic} 
and is helpful
in further minimising our core language. Expressions comprise unit (), 
constants \code{k}, 
variables \code{x}, 
%one-level field access \code{v.f} (rather than \code{v.f1.f2...} for the sake of simplicity), 
%data structure allocation \code{\textbf{new}\ d({v^*})},  
local variable declaration \code{\tau\ x;\ e}, method calls \code{mn({x^*}) }, 
variable assignments \code{x{:=}e}, 
%field update \code{v.f:=e}, 
expression sequences \code{e_1;e_2},
conditional expressions \code{{if}\ v\ {then}\ e_1\ {else}\ e_2},
event raises expression \code{ \textbf{event} [ev] },

@@

    \m{(Program)} &  \code{\mathcal{P} }&\code{  ::= }&
    \code{%{datat^*}  \  
   {meth^*}} 
\qquad   \qquad\quad 
    \m{(Basic\ Types)} \qquad  \code{\tau} \code{\quad  ::= \quad }
     \code{int\ |\ bool\ |\ void}
    \\

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{\tau \ mn  \ {( \tau\ x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ ()\ |\ k^\tau\ |\ x\ |\ 
    %v.f\ |\ 
    %\textbf{new}\ d({v^*}) \ |\ 
    \tau\ x;\ e \ |\ mn({x^*}) \ }
  \code{|\ x{:=}e\ |\ 
 % v.f:=e \ | 
 \ e_1;e_2\ 
  |\ {if}\ v\ {then}\ e_1\ {else}\ e_2 
 }  \\
 &   & & \code{
% |\ e_1}\ op\ \code{ e_2 \ 
| \ \textbf{event} [ev]\ 
  | \ \textbf{delay}\ [d] \ 
  | \ e_1\ \textbf{timeout}\  [d] \  e_2 \ 
  | \ \textbf{deadline}  \ [d] \ e \  
  }  \\
  ~\\

    \multicolumn{4}{c}{
  \code{k^\tau:constant\ of\ type\ \tau\ } 
  \qquad\quad
\code{    x, mn  ::\in \textbf{var}}  
\qquad\quad
 \code{(Events)  ev ::\in \Sigma 
\qquad\quad
\m{(Time\ Bounds)}\ \code{d} \code{\in \mathbb{Z^+}}    
}   
    }\\  
    \hline
  \end{array}  
  $
 \caption{A Core Imperative Language.} 
 \label{fig:code_language}
  \vspace{-1mm}
\end{figure}
}

A number of timed process constructs can be used to capture common 
real-time system behavior patterns. Without loss of generality, we 
assume d is an integer constant.  $\textbf{delay}\ [d]$ idles for exactly $d$ 
time units. In process $e_1\ \textbf{timeout}\ [d] e_2$, the first observable event of 
$e_1$ shall occur before d time units elapse (since the process starts). 
Otherwise, $e_2$ takes over control after exactly $d$ time units elapse. 
Process $\textbf{deadline}\ [d]\ e$ constrains $e$ to terminate before $d$ time units. 
In this setting, clock variables are 
made implicit and hence they cannot be compared with each other 
directly, which potentially allows efficient clock manipulation and 
hence system verification.

@@

\subsection{Semantics}
\label{subsec:Targetlanguage_Semantics}

>>>>

Here, \code{c} is for integer constants, \code{b} is for boolean values, 
 \code{mn} and \code{x} are for meta-variables.  
 \textbf{var} represents the countably infinite set of arbitrary distinct identifiers. 
\code{ev} refers to a singleton event coming from the finite 
set of events \code{\Sigma}. 
%Without loss of generality, any event can be further parametrized with 
%constant values, e.g. \code{\anyevent{a}.0} and  \code{\anyevent{a}}.1 are distinct two events. 
%leading to infinite alphabet.
We assume that programs we use are well-typed conforming to basic types \code{\tau} (we take () as the \code{void} type).
%Each events raising expression \code{\textbf{event}} is parametrized with an event, to trigger such a single event. 
A program \code{\mathcal{P}} comprises a list of 
global values initializations \code{(x:=v)^*} and 
a list of 
method declarations \code{{meth^*}}. 
Here, we use the \code{*} superscript to denote a finite list (possibly empty) of items, for example, \code{{x^*}} refers to a list of variables, \code{x_1, ...,\ x_n}. 
%Besides, \code{d} denotes the name of a user-defined data type, \code{f} denotes a field name.

Each method \code{meth} has a name \code{mn}, an expression-oriented body \code{e}, also is associated with a precondition  \code{\effect_{pre}} and a postcondition \code{\effect_{post}} (the syntax of effects specification \code{\effect} is given in \figref{fig:Syntax_of_Types_and_Effects}).
The language allows each iterative loop to be optimized to an equivalent 
tail-recursive method, where mutation on parameters is made visible 
to the caller. 
%The technique of translating away iterative loops is standard
%\citeref{insa2015automatic} 
%and is helpful
%in further minimising our core language.

Expressions comprise simple values \code{v};
a guarded process is written as \code{[b]e} where if b is true, 
then it behaves as \code{e}, else it idles until \code{b} becomes true; 
%one-level field access \code{v.f} (rather than \code{v.f1.f2...} for the sake of simplicity), 
%data structure allocation \code{\textbf{new}\ d({v^*})},  
%local variable declaration \code{\tau\ x;\ e}, 
method calls \code{mn({x^*})}; 
variable assignments \code{x{:=}e}; 
%field update \code{v.f:=e}, 
expression sequences \code{e_1;e_2};
parallel composition \code{e_1 || e_2}, where \code{e_1} and \code{e_2} 
may communicate via multi-party event synchronization or shared variables; 
conditional expressions \code{{if}\ v\ {then}\ e_1\ {else}\ e_2}.

@@

    \m{(Program)} &  \code{\mathcal{P} }&\code{  ::= }&
    \code{%{datat^*}  \  
   ((x:=v)^*, {meth^*})} 
%\qquad   \qquad\quad 
 %   \m{(Basic\ Types)} \qquad  \code{b} \code{\quad  ::= \quad }
  %   \code{int\ |\ bool\ |\ void}
    \\

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{ mn  \ {(x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Values)} & \code{v} &\code{  ::= } & ()\ |\ c \ | \ b \ |\ x\  
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ v \ | \ 
    [b]\ e \ |\ mn({v^*}) \ }
  \code{|\ x{:=}e\ |\ 
 % v.f:=e \ | 
 \ e_1;e_2\ | \ e_1 || e_2 \ 
  |\ {if}\ b\ {then}\ e_1\ {else}\ e_2 
 }  \\
 &   & & \code{
% |\ e_1}\ op\ \code{ e_2 \ 
| \ \textbf{event} [ev]\ 
  | \ \textbf{delay} [d] \ 
  | \ e_1\ \textbf{timeout} [d] \  e_2 \ 
  | \ \textbf{deadline}  [d] \ e \  
  }  \\
  ~\\

    \multicolumn{4}{c}{
 % \code{k^\tau:constant\ of\ type\ \tau\ } 
 % \qquad\quad
\code{c  \in \mathbb{Z}}
\qquad \ \ 
\code{b  \in \mathbb{B}}
\qquad \ \ 
\code{x, mn  ::\in \textbf{var}}  
\qquad\ \ 
 \code{(Events)  ev \in \Sigma 
\qquad\ \ 
\m{(Time\ Bounds)}\ \code{d} \code{\in \mathbb{N}}    
}   
    }\\  
    \hline
  \end{array}  
  $
 \caption{A Core First-Order Timed Imperative Language, \timedL.} 
 \label{fig:code_language}

\end{figure}
}

Apart for the event raising expression \code{ \textbf{event} [ev] }, 
a number of timed process constructs can be used to capture common 
real-time system behavior patterns. Without loss of generality, we 
assume d is a non-negative integer constant.  $\textbf{delay} [d]$ idles for exactly $d$ 
time units. In process $e_1\ \textbf{timeout} [d]\ e_2$, the first observable event of 
$e_1$ shall occur before d time units elapse (since the process starts). 
Otherwise, $e_2$ takes over control after exactly $d$ time units elapse. 
Process $\textbf{deadline} [d]\ e$ constrains $e$ to terminate before $d$ time units. 
In this setting, clock variables are 
made implicit and hence they cannot be compared with each other 
directly, which potentially allows efficient clock manipulation and 
hence system verification.

@@

\subsection{Semantics of the target language}
\label{subsec:Targetlanguage_Semantics}
